### 2017-03-29 intro.tex
Added Tessa's comments

\section{The Cosmological Context of Galaxy Formation} The modern picture of
galaxy formation sits within the context of a $\Lambda$ Cold Dark Matter
($\Lambda CDM$) cosmology \citep{Rees1977,White1978,Blumenthal1984,Mo1998}.
This is a universe which has a no curvature ($\Omega = 1$), and today has the
majority of its mass-energy in dark energy $\Lambda$ ($\Omega_\Lambda =
0.6911\pm0.0062$), and most of its matter in a collisionless form (dark matter)
($\Omega_m = 0.3089\pm0.0062$) \citep{Planck2015b}.  Roughly $5\%$ of the
matter-energy density of the universe exists as baryons ($\Omega_b =
0.048613\pm0.00031$).  The composition of the universe, and the precise value of
the mass-energy density for these components sets the lifetime of the universe,
how rapidly it expands (or contracts), and exactly determines the growth of
large-scale structure.  Within this structure is ultimately where galaxies will
form.  All of this began $13.799\pm0.021$ billion years ago \citep{Planck2015b},
with the universe emerging from a state of high temperature and density,
expanding to form what we see today.  The primordial elemental abundances, with
$\sim4/5$ of the universe's baryons in hydrogen, and the remainder in helium,
were established just a few minutes after the big bang, during a brief period of
``Big Bang Nucleosynthesis'' \citep{Alpher1948}.  The remaining heavier elements
have been formed via stellar nucleosynthesis, through fusion in the cores of
stars and in energetic processes that take place during supernovae (SN)
\citep{Wagoner1967}, or through the capture of high energy cosmic ray protons
and $\alpha$ particles by carbon, nitrogen, or oxygen nucleii
\citep{Reeves1970}.  The chemical composition of the universe establishes the
kinds of stars that form within galaxies and the rates at which gas within the
interstellar medium (ISM) and beyond cools.

Much of what we know about the cosmology of our universe comes from observations
of the first light to escape the universe's surface of last scattering as it
became optically thin: the cosmic microwave background radiation (CMB).  
When the CMB was emitted, the baryons of the universe were in the form of a
nearly isothermal ($\delta T/T\sim10^{-5}$) gas of hydrogen and helium.  Prior
to a redshift of $z\approx1100$, the universe had a high ionization fraction
such that the mean free path for photons was short, making all of space
effectively opaque.  Once the universe became cool enough that hydrogen became
fully neutral, the optical depth fell low enough that photons could free-stream,
decoupled from baryons.  We observe this ultraviolet light today, redshifted to
microwaves with a temperature of $2.729\pm0.004\;\rm{K}$ \citep{Fixsen1996}.
When the power spectrum of this radiation is measured \citep{Spergel2003},
features in that spectrum can be used to determine the components of the
universe (with some degeneracy that can be resolved via SN Ia and large scale
structure measurements
\citealt{Riess1998,Perlmutter1999,Beutler2011,Blake2011}). These features also
reveal the spectrum of density perturbations that will eventually form galaxies
\citep{Press1974,Peebles1980}.   The CMB is the single most important
observational tool for understanding the cosmology of our universe.

The formation of structure within the universe began with the gravitational
collapse of small, linear density perturbations seeded by quantum fluctuations
amplified by inflation \citep{Guth1981,Linde1982}.  These perturbations grow
through gravitational collapse.  Until $z\sim1100$, only dark matter
perturbations were able to grow.  Prior to this, baryons were coupled to
photons, causing density perturbations to oscillate as stable sound waves.  This
early collapse of matter meant that when decoupling occurred at $z\sim1100$,
baryons were able to collapse into the deeper potential wells grown out of
dark-matter density perturbations.  These overdensities expand slower that the
surrounding medium, as they experience a drag from their own self-gravity.
Eventually, these regions stop expanding and begin collapsing at the turnaround
time $t_{turn}$, and virialize at roughly $2\;t_{turn}$ \citep{Peebles1980}.
The first galaxies likely began to form around $z\sim20-50$, with minimum sizes
set by the baryonic physics of Silk damping and radiative cooling
\citep{Silk1968}.  Without the earlier growth of dark matter perturbations,
structure formation would have occurred much later, and the number of galaxies
and distribution of their masses would be different than those we see today
\citep{Davis1985}.

The density perturbation spectrum describes the complete statistics of the
Gaussian random field from which structure forms. \citet{Press1974} were able to
show this produces a halo mass spectrum at a given time with a form that matched
the functional form for the observed galaxy mass function
\citet{Schechter1976}:
\begin{equation}
    \frac{dn}{dM}(M,t) =
    \sqrt{\frac{2}{\pi}}\frac{\rho_0\delta_c(t)}{M^2\sigma(M)} 
    \left\lvert\frac{d\ln{\sigma}}{d\ln{M}}\right\rvert
    \exp{\left[-\frac{\delta^2_c(t)}{2\sigma^2(M)}\right]}
\end{equation}
In this equation, $\rho_0$ is the mean density of the universe, and
$\delta_c(t)$ is the critical overdensity for collapse at a given time $t$.  The
mass variance spectrum, $\sigma^2(M)$ can be determined observationally using
the CMB power spectrum \citep{Fixsen1996,Spergel2003,Planck2015} and an
assumption of a specific cosmology.  This function gives an approximate
distribution of masses for the dark matter halos that galaxies will begin to
form within.  As these halos collapse, non-linear gravitational tugs from their
surroundings will impart torques on them \citep{Barnes1987}, causing halos to
have a distribution of angular momenta.  High resolution N-body simulations have
found that dark matter halos have both a universal density profile
\citep{Navarro1996,Merritt2006} as well as a universal angular momentum profile
\citep{Bullock2001}.  These dark matter halos serve as the environment in
which the collapse of baryons and formation of stars take place, fed by
accretion along large-scale filaments.

@@

As gas does begin to cool within these halos, it will lose pressure support (by
radiating away thermal energy), but {\it not} angular momentum.  For large
spiral galaxies, this means that the baryons, which typically carry the same
specific angular momentum as the halo itself, will flatten to form an
exponential disc as they cool \citep{Fall1980,Mo1998}.    A combination
of heating and cooling processes will render this gas multiphase, containing
both warm/hot ($T>10^4\;\rm{K}$) and cold ($T\sim10-100\;\rm{K}$) gas in
quasistatic equilibrium \citep{McKee1977}.  This multiphase state is highly
dependent on both the metallicity and the local heating rates (which
themselves can vary with redshift) \citep{Field1969,Wolfire1995,Norman1997}.  In
the disc of the galaxy, this dense multiphase gas forms the ISM from which stars
collapse and are born.

The formation of stars has a dramatic impact on the galaxy.  As molecular clouds
begin to construct O/B stars with masses $>5\Msun$, these stars immediately
begin to obliterate the clouds from which they form.  A typical $10^5\Msun$
cluster will have an ionizing UV luminosity of $\sim10^5\Lsun$ during its first
few Myr of life.  This UV flux ionizes the surrounding gas, heating it to
$\sim10^4\;\rm{K}$ \citep{Dale2005,Murray2011}.  Along with ionizing radiation,
massive stars drive fast $v>10^3\kms$ stellar winds, further shocking and
destroying the cold, dense gas which stars form from.  By the time the most
massive stars reach the end of their short lives, $\sim4$ Myr later, they
detonate as core-collapse supernovae not in dense ($n>100\hcc$) clouds, but in
their significantly more diffuse remnants \citep{Rogers2013}.  The longer
cooling times this results in allows supernovae to expand through the ISM,
driving turbulance \citep{Ostriker2010} and filling most of the volume with hot,
diffuse gas.  Much of this hot gas is in fact able to leave the ISM, driven out
as hot, buoyant, high-entropy gas.  These feedback-driven outflows remove gas
from the disc, slowing the rate of star formation, and polluting the
intergalactic medium with metals from stellar ejecta.  In the most massive
galaxies ($M>10^{12}\Msun$), these outflows can be driven by the accretion
luminosity of the central supermassive black hole (SMBH).  These Active Galactic
Nuclei (AGN) are observed in giant elliptical galaxies, and can be seen blowing
hot bubbles within the intracluster medium of nearby galaxy clusters.  Further
discussion on feedback can be found in section~\ref{intro:fb}.

@@

These are of course all highly idealized, simple models for the broadest strokes
of galaxy formation.  Each of these halos will begin to undergo mergers and
interactions \citet{Navarro1993,Kauffmann1999,Cole2000} with each other that can
strip stars and gas through tidal and ram pressure forces
\citep{Gunn1972,Dressler1980}.  Inside each of these halos, the process of disc
collapse, the formation of a multiphase ISM, star formation, and feedback from
those stars involves the complex, nonlinear interplay of radiation,
hydrodynamics, magnetic fields, and gravity on length scales as much as
$10^{11}$ times smaller than the virial radius.  None of this physics is simple
enough to be fully captured by analytic models. As figure 1.1 shows, the
observed mass function for galaxies does not match what would be expected from
the total baryon budget for each halo, implying that at least some of these
processes must change the galactic star formation efficiency as a function of
halo mass, establishing the population of galaxies we see today.  Numerical
simulations have therefore become a keystone for galaxy formation research.

\section{Numerical Simulations of Galaxies}
N-body Simulations of galaxies actually predate {\it digital} simulations of
galaxies.  \citet{Holmberg1941} presented, as the author termed it, a ``new
integration procedure'' that was the first attempt to simulate the
evolution of a galaxy.  By relying on the fact that light flux, just as gravity,
scales as $r^{-2}$, Holmberg was able to construct a simulated galaxy disc out
of light bulbs, and then using photocells to measure the flux at each bulb's
position, repositioned the bulbs using the forces ``simulated'' by the
light-measuring process.  This allowed him to painstakingly trace the tidal
interactions of passing galaxies as they merged with each other.

In the intervening 75 years, we have migrated to much less labor-intensive
simulations, relying on the constantly growing power of digital computers.  As
Gordon Moore first observed in 1965, the processing power of computers has been
continuously growing at an exponential rate for more than half a century.  This
has meant that, even without advances in numerical methods and algorithm design,
the ability of simulators today to model a system in high resolution is
monumentally greater than just 10 years ago.  Beyond the simple improvement of
hardware, significant advances have been made in the underlying algorithms used
in simulations, allowing the same hardware to perform larger, higher resolution
simulations than ever before.  These algorithmic improvements have increased the
speed and scalability of simulation codes even faster than improvements in
hardware.

@@

\subsection{Gravity}
Gravity is typically solved using algorithms designed to minimize the workload
that would be needed for a direct summation of each particle or grid cell
($\mathcal{O}(N^2)$ for N grid cells/particles).  This becomes intractable for
larger numbers of resolution elements, as the workload will scale as the sixth
power of the linear resolution (an order of magnitude better linear resolution
will be $10^6$ times more expensive!).  In order to reduce these costs,
algorithms have been developed that allow for more efficient gravitational
force calculations.  Tree-based methods \citep{Barnes1986} reduce the
computational cost of long-ranged (and thus less significant) force
calculations by using space-partitioning trees.  Short-range force calculations
are done particle-by-particle, while long-range forces are calculated against
larger tree cells, thus reducing the cost of calculating gravity to
$\mathcal{O}(N\log N)$.  Particle-Particle/Particle-Mesh ($P^3M$) methods
\citep{Couchman1991} produce similar scalings, but rather than using a
spatial tree rely on a spectral Fourier-based approach to calculate long-ranged
interactions.  Recently, the Fast Multipole Method \citep{Greengard1987} has
been applied to the calculation of gravitational forces
\citep{Dehnen2002,Hahn2013} using a multipole expansion of Green's functions.
This allows for extremely efficient gravity calculation, scaling linearly with
particle number ($\mathcal{O}(N)$).  This method allows it to now be possible to
run N-body simulations with greater than $10^{12}$ particles\citep{Potter2016}.

\subsection{Hydrodynamics}
Hydrodynamical solvers usually come in one of two classes: Eulerian or
Lagrangian.  Eulerian methods discretize space into fixed volume elements with
fixed boundaries.  By breaking a simulation volume into a grid of cells that
each contains scalar and vector quantities those quantities can be advected
between cells according to the solver scheme.  Eulerian schemes typically allow
for higher-order solvers, lower intrinsic noise, and better resolved shocks
(when at rest relative to the grid) \citep{Teyssier2002,Stone2008,Bryan2014}.
They suffer from intrinsic mixing and diffusion that scales with the flow
velocity relative to the grid, poor integration of circular orbits, and costly
timestep limits \citep{Agertz2007,Tasker2008}.  A variety of Eulerian solvers
exist, ranging from the staggered mesh, high-order upwind \& artificial
viscosity scheme ZEUS, used in codes such as {\sc ENZO}
\citep{Stone1999,Bryan2014}, to the 2nd order Godunov method MUSCL used by {\sc
ramses} \citep{vanLeer1976,Teyssier2002}, or the 3rd order PPM scheme of {\sc
Athena} \citep{Stone2008}.  Lagrangian methods instead discretize mass, breaking
the volume into discrete points, each of which moves at the fluid
velocity to advect the quantities it contains, with smoothing functions used to
interpolate values in between particles; hence the name for the most common
Lagrangian method, Smoothed Particle Hydrodynamics (SPH).  Lagrangian methods
couple easily to tree and particle-mesh codes for gravity calculation, are
Galilean invariant, and intrinsically conserve mass, energy, and momentum
(linear and angular) \citep{Katz1996,Wadsley2004,Springel2005}.  The perfect
conservation of angular momentum is critical for accurately tracking orbits. SPH
methods typically use artificial viscosity to handle shocks, and as such have
difficulty in capturing subsonic turbulence \citep{Bauer2012}.  Recently, new
methods have begin to extend the basic SPH method using higher order gradient
estimates that are allowing accuracy on subsonic tests that rival the best
Eulerian schemes \citep{Springel2010,Hopkins2015,Valdarnini2016}.

Modern galaxy simulations also now use a variety of strategies for judiciously
applying computational power where it is most needed.  Eulerian methods will
commonly use \textit{Adaptive Mesh Refinement} to increase resolution in regions
of interest (typically where gas is densest).  Lagrangian methods automatically
do this, as a denser region definitionally must contain more fluid elements.

\subsection{Designing Simulations}
Beyond the choice of algorithms, the construction of appropriate inital
conditions is also important.  To study large populations of galaxies, along
with large-scale structure, large-volume cosmological boxes with volumes
$\gtrsim 10^6\;\rm{Mpc^3}$, but with spatial resolutions $\sim \rm{kpc}$ are
used (for example, the {\it Illustris} simulations have a softening length of
$1.4\;rm{kpc}$ \citealt{Vogelsberger2014}).  Studying how individual galaxies
form is typically done with cosmological zoom simulations, where a comparatively
cheap dark-matter only simulation is first run to select a region of interest,
which is then resimulated with higher resolution and baryonic physics included
\citep{Navarro1993}. Zoom-in simulations can typically achieve an order of
magnitude or better resolution than full cosmological boxes.  A number of groups
today are using these zoom-in simulations to study the impact of feedback
processes over cosmic time \citep{Stinson2013,Hopkins2014,Agertz2015}.  Isolated
galaxy disc simulations use a constructed disc profiles containing gas and
stars, embedded within a dark halo. These cannot show how galaxies form in a
cosmological environment, but can be run with $\sim \rm{pc}$ resolution to study
closely how internal processes within a galaxy take place
\citep{Hopkins2011,Benincasa2016}.  Details of these internal processes are
discussed in section 1.3.

The earliest attempt to track the formation of a galaxy in a cosmological
context with the effects of star formation and feedback included was by
\citet{Katz1992}.  This simulation used the Lagrangian TREESPH 
\citep{Hernquist1989} to evolve a volume of gas and dark matter from an initial
linear perturbation, including small-scale power derived from CDM predictions
\citep{Zeldovich1970,Peebles1982}.  Gas was allowed to cool radiatively using
rates calculated for primordial gas.  This study introduced a model for star
formation that is still the standard for most simulations of galaxy formation.
Observations \citep{Kennicutt1998} have suggested that the star formation rate
is a power law of the gas density: $\dot\Sigma_*\propto\Sigma_{gas}^\alpha$, with
$\alpha=1-2$.  If it is assumed that gas collapses to form stars in some
multiple $c_*$ of its freefall time \citep{Schmidt1959}, $t_{ff}$, a simple
3-dimensional star formation model can be built with the form:
\begin{equation}
    \dot\rho_* = \frac{c_*\rho_{gas}}{t_{ff}} = c_*\sqrt{\frac{\rho^3_{gas}}{4\pi G}}
\end{equation}
\citet{Katz1992} used a $c_*$ parameter of $0.1$, which is within the range of
what is typically used in simulations today ($1-100\%$).  On top of the model
for star formation, feedback was included without the use of any significant
subgrid model:  $10^{51}\;\rm{erg/SN}$ worth of energy was injected back into
the ISM by smoothing it over a number of resolution elements.  One of the key
findings of \citet{Katz1992} was that the vast majority of SN energy was
radiated away, making them very ineffective at regulating star formation within
the galaxy.  In the end, \citet{Katz1992} found that the peak star formation in
the simulated galaxy was nearly 10 times greater than in observed galaxies.
\citet{Thacker2000} later showed that the smoothing of energy used by
\citet{Katz1992} resulted in a large overestimation of the cooling rates in
feedback-heated gas. Since then, much work has gone into developing models for
feedback that address the issue of overcooling that was present in these
earlier simulations.  These efforts will be described in detail in the next
section.

>>>>

\section{The Cosmological Context of Galaxy Formation} 
The modern picture of galaxy formation sits within the context of a $\Lambda$
Cold Dark Matter ($\Lambda CDM$) cosmology
\citep{Rees1977,White1978,Blumenthal1984,Mo1998}.  This is a universe which has
a no curvature ($\Omega = 1$), and today has the majority of its mass-energy in
dark energy $\Lambda$ ($\Omega_\Lambda = 0.6911\pm0.0062$), and most of its
matter in a collisionless form (dark matter) ($\Omega_m = 0.3089\pm0.0062$)
\citep{Planck2015b}.  Roughly $5\%$ of the matter-energy density of the universe
exists as baryons ($\Omega_b = 0.048613\pm0.00031$).  The composition of the
universe, and the precise value of the mass-energy density for these components,
sets the lifetime of the universe, how rapidly it expands (or contracts), and
exactly determines the growth of large-scale structure.  Within this structure
is ultimately where galaxies will form.  All of this began $13.799\pm0.021$
billion years ago \citep{Planck2015b}, with the universe emerging from a state
of high temperature and density, expanding to form what we see today.  The
primordial elemental abundances, with $\sim4/5$ of the universe's baryons in
hydrogen (the remainder in helium), were established just a few minutes after
the big bang, during a brief period of ``Big Bang Nucleosynthesis''
\citep{Alpher1948}.  The remaining heavier elements have been formed via stellar
nucleosynthesis, through fusion in the cores of stars and in energetic processes
that take place during supernovae (SN) \citep{Wagoner1967}. A small fraction of
these metals form via the capture of high energy cosmic ray protons and $\alpha$
particles by carbon, nitrogen, or oxygen nucleii \citep{Reeves1970}.  The
chemical composition of the universe establishes the kinds of stars that form
within galaxies and the rates at which gas within the interstellar medium (ISM)
and beyond cools.

Much of what we know about the cosmology of our universe comes from observations
of the first light to escape the universe's surface of last scattering as it
became optically thin: the cosmic microwave background radiation (CMB).  
When the CMB was emitted, the baryons of the universe were in the form of a
nearly isothermal ($\delta T/T\sim10^{-5}$) gas of hydrogen and helium.  Prior
to a redshift of $z\approx1100$, the universe had a high ionization fraction
such that the mean free path for photons was short, making all of space
effectively opaque.  Once the universe became cool enough that hydrogen became
fully neutral, the optical depth fell low enough that photons could free-stream,
decoupled from baryons.  We observe this ultraviolet light today, redshifted to
microwaves with a temperature of $2.729\pm0.004\;\rm{K}$ \citep{Fixsen1996}.
When the power spectrum of this radiation is measured \citep{Spergel2003},
features in that spectrum can be used to determine the components of the
universe (with some degeneracy that can be resolved via SN Ia and large scale
structure measurements
\citep{Riess1998,Perlmutter1999,Beutler2011,Blake2011}). These features also
reveal the spectrum of density perturbations that will eventually form galaxies
\citep{Press1974,Peebles1980}.   The CMB is the single most important
observational tool for understanding the cosmology of our universe.

The formation of structure within the universe began with the gravitational
collapse of small, linear density perturbations seeded by quantum fluctuations
amplified by inflation \citep{Guth1981,Linde1982}.  These perturbations grow
through gravitational collapse.  Until $z\sim1100$, only dark matter
perturbations were able to grow.  Prior to this, baryons were coupled to
photons, causing density perturbations to oscillate as stable sound waves.  This
early collapse of matter meant that when decoupling occurred at $z\sim1100$,
baryons were able to collapse into the deeper potential wells grown out of
dark matter density perturbations.  These overdensities expand slower that the
surrounding medium, as they experience a drag from their own self-gravity.
Eventually, these regions stop expanding and begin collapsing at the turnaround
time $t_{turn}$, and virialize at roughly $2\;t_{turn}$ \citep{Peebles1980}.
The first galaxies likely began to form around $z\sim20-50$, with minimum sizes
set by the baryonic physics of Silk damping and radiative cooling
\citep{Silk1968}.  Without the earlier growth of dark matter perturbations,
structure formation would have occurred much later, and the number of galaxies
and distribution of their masses would be different than those we see today
\citep{Davis1985}.

The density perturbation spectrum describes the complete statistics of the
Gaussian random field from which structure forms. \citet{Press1974} were able to
show this produces a halo mass spectrum at a given time with a form that matched
the functional form for the observed galaxy mass function
\citep{Schechter1976}:
\begin{equation}
    \frac{dn}{dM}(M,t) =
    \sqrt{\frac{2}{\pi}}\frac{\rho_0\delta_c(t)}{M^2\sigma(M)} 
    \left\lvert\frac{d\ln{\sigma}}{d\ln{M}}\right\rvert
    \exp{\left[-\frac{\delta^2_c(t)}{2\sigma^2(M)}\right]}
\end{equation}
In this equation, $\rho_0$ is the mean density of the universe, and
$\delta_c(t)$ is the critical overdensity for collapse at a given time $t$.  The
mass variance spectrum, $\sigma^2(M)$ can be determined observationally using
the CMB power spectrum \citep{Fixsen1996,Spergel2003,Planck2015} and an
assumption of a specific cosmology.  This function gives an approximate
distribution of masses for the dark matter halos in which galaxies will begin to
form.  As these halos collapse, non-linear gravitational tugs from their
surroundings will impart torques on them \citep{Barnes1987}, causing halos to
have a distribution of angular momenta.  High resolution N-body simulations have
found that dark matter halos have both a universal density profile
\citep{Navarro1996,Merritt2006} as well as a universal angular momentum profile
\citep{Bullock2001}.  These dark matter halos serve as the environment in
which the collapse of baryons and formation of stars take place, fed by
accretion along large-scale filaments.

@@

As gas begins to cool within these halos, it will lose pressure support (by
radiating away thermal energy), but {\it not} angular momentum.  For large
spiral galaxies, this means that the baryons, which typically carry the same
specific angular momentum as the halo itself, will flatten to form an
exponential disc as they cool \citep{Fall1980,Mo1998}.    A combination
of heating and cooling processes will render this gas multiphase, containing
both warm/hot ($T>10^4\;\rm{K}$) and cold ($T\sim10-100\;\rm{K}$) gas in
quasistatic equilibrium \citep{McKee1977}.  This multiphase state is highly
dependent on both the metallicity and the local heating rates (which
themselves can vary with redshift) \citep{Field1969,Wolfire1995,Norman1997}.  In
the disc of the galaxy, this dense multiphase gas forms the ISM from which stars
collapse and are born.

The formation of stars has a dramatic impact on the galaxy.  As molecular clouds
begin to construct O/B stars with masses $>5\Msun$, these stars immediately
begin to obliterate the clouds from which they form.  A typical $10^5\Msun$
cluster will have an ionizing UV luminosity of $\sim10^5\Lsun$ during its first
few Myr of life.  This UV flux ionizes the surrounding gas, heating it to
$\sim10^4\;\rm{K}$ \citep{Dale2005,Murray2011}.  Along with ionizing radiation,
massive stars drive fast $v>10^3\kms$ stellar winds, further shocking and
destroying the cold, dense gas which stars form from.  By the time the most
massive stars reach the end of their short lives, $\sim4$ Myr later, they
detonate as core-collapse supernovae not in dense ($n>100\hcc$) clouds, but in
their significantly more diffuse remnants \citep{Rogers2013}.  The longer
cooling times this results in allow supernovae to expand through the ISM,
driving turbulance \citep{Ostriker2010} and filling most of the volume with hot,
diffuse gas.  Much of this hot gas is in fact able to leave the ISM, driven out
as buoyant, high-entropy gas.  These feedback-driven outflows remove gas
from the disc, slowing the rate of star formation, and polluting the
intergalactic medium with metals from stellar ejecta.  In the most massive
galaxies ($M>10^{12}\Msun$), these outflows can be driven by the accretion
luminosity of the central supermassive black hole (SMBH).  These active galactic
nuclei (AGN) are observed in giant elliptical galaxies, and can be seen blowing
hot bubbles within the intracluster medium of nearby galaxy clusters.  Further
discussion on feedback can be found in section~\ref{intro:fb}.

@@

These are of course all highly idealized, simple models for the broadest strokes
of galaxy formation.  Each of these halos will begin to undergo mergers and
interactions \citep{Navarro1993,Kauffmann1999,Cole2000} with each other that can
strip stars and gas through tidal and ram pressure forces
\citep{Gunn1972,Dressler1980}.  Inside each of these halos, the process of disc
collapse, the formation of a multiphase ISM, star formation, and feedback from
those stars involves the complex, nonlinear interplay of radiation,
hydrodynamics, magnetic fields, and gravity on length scales as much as
$10^{11}$ times smaller than the virial radius.  None of this physics is simple
enough to be fully captured by analytic models. As figure 1.1 shows, the
observed mass function for galaxies does not match what would be expected from
the total baryon budget for each halo, implying that at least some of these
processes must change the galactic star formation efficiency as a function of
halo mass, establishing the population of galaxies we see today.  Numerical
simulations have therefore become a keystone for galaxy formation research.

\section{Numerical Simulations of Galaxies}
N-body simulations of galaxies actually predate {\it digital} simulations of
galaxies.  \citet{Holmberg1941} presented, as the author termed it, a ``new
integration procedure'' that was the first attempt to simulate the
evolution of a galaxy.  By relying on the fact that light flux, just as gravity,
scales as $r^{-2}$, Holmberg was able to construct a simulated galaxy disc out
of light bulbs, and then using photocells to measure the flux at each bulb's
position, repositioned the bulbs using the forces ``simulated'' by the
light-measuring process.  This allowed him to painstakingly trace the tidal
interactions of passing galaxies as they merged with each other.

In the intervening 75 years, we have migrated to much less labor-intensive
simulations, relying on the constantly growing power of digital computers.  As
Gordon Moore first observed in 1965, the processing power of computers has been
continuously growing at an exponential rate for more than half a century.  This
has meant that, even without advances in numerical methods and algorithm design,
the ability of simulators today to model a system in high resolution is
monumentally greater than just 10 years ago.  Beyond the simple improvement of
hardware, significant advances have been made in the underlying algorithms used
in simulations, allowing the same hardware to perform larger, higher-resolution
simulations than ever before.  These algorithmic improvements have increased the
speed and scalability of simulation codes even faster than improvements in
hardware.

@@

\subsection{Gravity}
Gravity is typically solved using algorithms designed to minimize the workload
that would be needed for a direct summation of each particle or grid cell
($\mathcal{O}(N^2)$ for N grid cells/particles).  This becomes intractable for
larger numbers of resolution elements, as the workload will scale as the sixth
power of the linear resolution (an order of magnitude better linear resolution
will be $10^6$ times more expensive!).  In order to reduce these costs,
algorithms have been developed that allow for more efficient gravitational
force calculations.  Tree-based methods \citep{Barnes1986} reduce the
computational cost of long-ranged (and thus less significant) force
calculations by using space-partitioning trees.  Short-range force calculations
are done particle-by-particle, while long-range forces are calculated against
larger tree cells, thus reducing the cost of calculating gravity to
$\mathcal{O}(N\log N)$.  Particle-Particle/Particle-Mesh ($P^3M$) methods
\citep{Couchman1991} produce similar scalings, but rather than using a
spatial tree, rely on a spectral Fourier-based approach to calculate long-ranged
interactions.  Recently, the Fast Multipole Method \citep{Greengard1987} has
been applied to the calculation of gravitational forces
\citep{Dehnen2002,Hahn2013} using a multipole expansion of Green's functions.
This allows for extremely efficient gravity calculation, scaling linearly with
particle number ($\mathcal{O}(N)$).  This method allows it to now be possible to
run N-body simulations with greater than $10^{12}$ particles \citep{Potter2016}.

\subsection{Hydrodynamics}
Hydrodynamical solvers usually come in one of two classes: Eulerian or
Lagrangian.  Eulerian methods discretize space into fixed-volume elements with
fixed boundaries.  By breaking a simulation volume into a grid of cells that
each contains scalar and vector quantities, those quantities can be advected
between cells according to the solver scheme.  Eulerian schemes typically allow
for higher-order solvers, lower intrinsic noise, and better resolved shocks
(when at rest relative to the grid) \citep{Teyssier2002,Stone2008,Bryan2014}.
They suffer from intrinsic mixing and diffusion that scales with the flow
velocity relative to the grid, poor integration of circular orbits, and costly
timestep limits \citep{Agertz2007,Tasker2008}.  A variety of Eulerian solvers
exist, ranging from the staggered mesh, high-order upwind \& artificial
viscosity scheme ZEUS, used in codes such as {\sc ENZO}
\citep{Stone1999,Bryan2014}, to the 2nd order Godunov method MUSCL used by {\sc
ramses} \citep{vanLeer1976,Teyssier2002}, or the 3rd order PPM scheme of {\sc
Athena} \citep{Stone2008}.  Lagrangian methods instead discretize mass, breaking
the volume into discrete points, each of which moves at the fluid
velocity to advect the quantities it contains, with smoothing functions used to
interpolate values in-between particles; hence the name for the most common
Lagrangian method, Smoothed Particle Hydrodynamics (SPH).  Lagrangian methods
couple easily to tree and particle-mesh codes for gravity calculation, are
Galilean invariant, and intrinsically conserve mass, energy, and momentum
(linear and angular) \citep{Katz1996,Wadsley2004,Springel2005}.  The perfect
conservation of angular momentum is critical for accurately tracking orbits. SPH
methods typically use artificial viscosity to handle shocks, and as such have
difficulty with capturing subsonic turbulence \citep{Bauer2012}.  Recently, new
methods have begun to extend the basic SPH method using higher order gradient
estimates that are allowing accuracy on subsonic tests that rival the best
Eulerian schemes \citep{Springel2010,Hopkins2015,Valdarnini2016}.

Modern galaxy simulations also now use a variety of strategies for judiciously
applying computational power where it is most needed.  Eulerian methods will
commonly use Adaptive Mesh Refinement (AMR)(AMR)  xies to increase resolution in
regions of interest (typically where gas is densest).  Lagrangian methods
automatically do this, as a denser region definitionally must contain more fluid
elements.

\subsection{Designing Simulations}
Beyond the choice of algorithms, the construction of appropriate inital
conditions is also important.  To study large populations of galaxies (along
with large-scale structure) large-volume cosmological boxes with volumes
$\gtrsim 10^6\;\rm{Mpc^3}$, and with spatial resolutions $\sim \rm{kpc}$ are
used (for example, the {\it Illustris} simulations have a softening length of
$1.4\;\rm{kpc}$ \citep{Vogelsberger2014b}).  Studying how individual galaxies
form is typically done with cosmological zoom simulations, where a comparatively
cheap dark-matter only simulation is first run to select a region of interest,
which is then resimulated with higher resolution and baryonic physics included
\citep{Navarro1993}. Zoom-in simulations can typically achieve an order of
magnitude or better resolution than full cosmological boxes.  A number of groups
today are using these zoom-in simulations to study the impact of feedback
processes over cosmic time \citep{Stinson2013,Hopkins2014,Agertz2015}.  Isolated
galaxy disc simulations use constructed disc profiles containing gas and
stars, embedded within a dark halo. These cannot show how galaxies form in a
cosmological environment, but can be run with $\sim \rm{pc}$ resolution to study
closely how internal processes within a galaxy take place
\citep{Hopkins2011,Benincasa2016}.  Details of these internal processes are
discussed in section 1.3.

The earliest attempt to track the formation of a galaxy in a cosmological
context with the effects of star formation and feedback included was by
\citet{Katz1992}.  This simulation used the Lagrangian TREESPH 
\citep{Hernquist1989} to evolve a volume of gas and dark matter from an initial
linear perturbation, including small-scale power derived from CDM predictions
\citep{Zeldovich1970,Peebles1982}.  Gas was allowed to cool radiatively using
rates calculated for primordial gas.  This study introduced a model for star
formation that is still the standard for most simulations of galaxy formation.
Observations \citep{Kennicutt1998} have suggested that the star formation rate
is a power law of the gas density: $\dot\Sigma_*\propto\Sigma_{gas}^\alpha$, with
$\alpha=1-2$.  If it is assumed that gas collapses to form stars in some
multiple $c_*$ of its freefall time \citep{Schmidt1959}, $t_{ff}$, a simple
3-dimensional star formation model, can be built with the form:
\begin{equation}
    \dot\rho_* = \frac{c_*\rho_{gas}}{t_{ff}} = c_*\sqrt{\frac{\rho^3_{gas}}{4\pi G}}
\end{equation}
\citet{Katz1992} used a $c_*$ parameter of $0.1$, which is within the range of
what is typically used in simulations today ($1-100\%$).  On top of the model
for star formation, feedback was included without the use of any significant
subgrid model:  $10^{51}\;\rm{erg/SN}$ worth of energy was injected back into
the ISM by smoothing it over a number of resolution elements.  One of the key
findings of \citet{Katz1992} was that the vast majority of SN energy was
radiated away, making them very ineffective at regulating star formation within
the galaxy.  In the end, \citet{Katz1992} found that the peak star formation in
the simulated galaxy was nearly 10 times greater than in observed galaxies.
\citet{Thacker2000} later showed that the smoothing of energy used by
\citet{Katz1992} resulted in a large overestimation of the cooling rates in
feedback-heated gas. Since then, much work has gone into developing models for
feedback that address the issue of overcooling that was present in these
earlier simulations.  These efforts will be described in detail in the next
section.
